Filamentary transport in high-power H-mode conditions and in no/small-ELM regimes to predict heat and particle loads on PFCs for future devices

### 2016-11-15 Presentation/gpm_november2016_vianello.tex
Added last correction from Jens to EPM presentation

\begin{frame}{L-Mode studies: AUG/2}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item<1-> Profile modified by an increase of blob-size and change of
        \alert{packing fraction: $f_{fil} = \nu_{fil}\tau_{AC}$}
        \onslide<2->{\textcolor{blue}{and filament relative density}
          {\footnotesize (Carralero 2016 in preparation)}}
      \item<3-> As a consequence the contribution of filaments to
        radial transport increases
      \item<4|only@4> \alert{Beware the change of frequency may be due to a
        modification of fluctuation velocity which is know to vary
        with densities ad normalized greenwald fraction \parencite{0029-5515-51-5-053020}}  
      \item<5> Parallel flow is strongly reduced whenever we increase
        the divertor collisionality
      \end{itemize}
    \end{column}
      \begin{column}{0.5\textwidth}
        \includegraphics<1>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16a}
        \includegraphics<2>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16b}
        \includegraphics<3>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16c}
        \includegraphics<4>[width=\textwidth]{../pdfbox/KoM15Nov/AgostiniNF2011}
        \includegraphics<5>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16d}
      \end{column}
    \end{columns}
  \end{frame}

@@

  \section{Open issues}
  \begin{frame}{Open and unresolved issues}
  \setbeamercovered{transparent}
  \begin{enumerate}[<+(1) | invisible@-+>]
    \item Does $\Lambda_{div}$ unify the picture among the devices?
      \alert{No, deos neutral pressure play a role?}
    \item If neutrals play a role is it at the divertor and/or at the midplane?
      \alert{Contradictory results if one include experiments with
        midplane puffing, JET and EIRENE simulations}
    \item Does L$_\parallel$ play any role? \alert{So far TCV suggests no
      dependence from f$_x$ scan but both TCV and MAST observe an
      I$_p$ dependence. MAST clearly state filament $\sigma_{\perp}$
      increases with L$_{\parallel}$ }
    \item Is cooling the divertor with fueling or with equivalent? \alert{Contradictory observations in AUG and JET}
    \item Do we observe same behavior in L and H-Mode? \alert{So
      far no as shown in H-Mode AUG. We need higher detachment
      condition and we need enough fueling}
      \item Is the density shoulder accompanied with an ion temperature shoulder?
%      \item What are the underlying mechanisms? 
    \end{enumerate}
  \end{frame}

@@

\begin{frame}{Focal points}
\begin{itemize}
	\item MST1 uniquely facilitates cross comparison between machines
	\item Several of the proposed experiments overlap and will be combined
	\item A cross machine experiment has the makings of settling open issues
	\item Therefore, we will allocate shots for cross-machine comparison
	\item ....but also to other proposed experiments
	\item Cross machine L-mode experiments: 
	\begin{enumerate}
		\item Investigate the role of neutrals 
		\item $I_p$ and $q_{95}$ scans
	\end{enumerate}
	\item Cross machine H-mode experiments. 
\end{itemize}	
\end{frame}

\begin{frame}{L-mode experiments}
	\begin{itemize}
		\item The role of neutrals in the shoulder formation is not understood
                  {\scriptsize(proponents: Carralero (\# 7), Militello
                    (\# 1, 2), Vianello (\# 13), Walkden (\# 9))}
		\begin{enumerate}
		\item We envisage to measure neutral gas profiles at the outboard midplane
                  using fast cameras with D$_{\alpha}$ filter, density
                  and temperature SOL profiles and appropriate code
                  (e.g. KN1D as in \parencite{Lipschultz:2005gg,Lipschultz:2016vk})
		\item Reciprocating probes and profiles available on
                  all machines
                \item Investigate role of fueling location (MAST)
		\end{enumerate}
		\item Disentangle the roles of $I_p$, $q_{95}$, and $L_{\|}$.
                  {\scriptsize(proponents: Carralero , Militello, Vianello, Tsui, WalkdenCarralero (\# 7), Militello
                    (\# 1, 2), Vianello (\# 13), Walkden (\# 9), Tsui
                    (\# 15))}
		\begin{enumerate}
			\item Carry out parameter scans on all machines
			\item Strive after similar machine configurations
		\end{enumerate}
	\end{itemize}
\end{frame}

@@

\begin{frame}{COMPASS collaboration}
\begin{itemize}
\item COMPASS is not an MST1 device, therefore no operational cost
  will be provided by MST1
\item However GA agree that MST1 can provide part of the ppy for
  running experiments in line with MST1 topics
\item Proposal \# 14 is focused filaments dynamics and shoulder
  formation both in L and H-Mode
\item Experimental plan can be therefore agreed in order to perform
  similar experiments (e.g. I$_p$ and $q_{95}$ scan in L-Mode)
  ensuring proper comparison between devices
\item COMPASS suppose 3ppy for running and evaluation of experiment
  pertaining Topic-21, among Czeck and other european laboratories.
\end{itemize}
\end{frame}

\begin{frame}{There are urgent experiment which cannot be addressed}
\begin{itemize}
	\item Detailed ion temp measurements. Must eventually be
          completed in internal programs. Are you willing to share the
          information among Topic 21 even though done internally?
	\item Topology investigations? For example the comparison
          USN/LSN/DN is not possible. Upper divertor information is
          actually less detailed. Better to use reverse B$_t$ or ion
          $\mathbf{B}\times\nabla\mathbf{B}$ pointing away from
          X-Point which is know to have influences in detachment as
          well \parencite{McLean}
	\item Strikepoint sweeping and cryo pump
	\item Parameters scans in H-mode
	\item X-point probes
\end{itemize}
\end{frame}

\begin{frame}{Modeling require}
  \begin{enumerate}
    \item Mean field modelling available for AUG (EMC3-EIRENCE) and
      MAST-U (SOLPS). So far we lack investigation for these scenario
      targeting TCV. Manpower needed from edge and SOL modelling
    \item Turbulence code: we should need self-consistent interaction
      with neutrals. Possible tools \texttt{BOUT ++}, \texttt{HESEL},
      \texttt{GBS}. Some of them can be run by people who will, very
      likely, apply for this topic, some other not
  \end{enumerate}
\end{frame}

\begin{frame}{Future and Miscellaneous}
	\begin{itemize}
		\item Possibility to piggybag on Quescient/small ELM H-mode shots (Topic 5,6, 18)
		\item Shortly after this meeting we will organize a follow-up meeting to reconcile ideas
                  from this discussion (date not fixed yet)
		\item All information will be gathered on the wiki
                \item Other channels for discussion/sharing. Slack
                  (mst1-topic21.slack.com), Github for code sharing
		\item We aim to arrive to a proper shared database of
                  filaments between the machines, where properties
                  will be obtained using agreed approaches so that a
                  cross-machine comparison paper could at the end be
                  finalized at a certain point		
	\end{itemize}
\end{frame}

\begin{frame}{Discussion agenda}
	\begin{enumerate}
		\item Priorities. These are our personal ideas. Do we agree? 
		\item \underline{Common measurement techniques}. Filament
                  properties, $\Lambda_{div}$ estimate and location,
                  Neutrals, \ldots 
 	\end{enumerate}
\end{frame}

>>>>

\begin{frame}{L-Mode studies: AUG/2}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item<1-> Profile modified by an increase of blob-size and change of
        \alert{packing fraction: $f_{fil} = \nu_{fil}\tau_{AC}$}
        \onslide<2->{\textcolor{blue}{and filament relative density}
          {\footnotesize (Carralero 2016 in preparation)}}
      \item<3-> As a consequence the contribution of filaments to
        radial transport increases
      \item<4|only@4> \alert{Beware the change of frequency may be due to a
        modification of fluctuation velocity which is known to vary
        with densities ad normalized greenwald fraction \parencite{0029-5515-51-5-053020}}  
      \item<5> Parallel flow is strongly reduced whenever we increase
        the divertor collisionality
      \end{itemize}
    \end{column}
      \begin{column}{0.5\textwidth}
        \includegraphics<1>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16a}
        \includegraphics<2>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16b}
        \includegraphics<3>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16c}
        \includegraphics<4>[width=\textwidth]{../pdfbox/KoM15Nov/AgostiniNF2011}
        \includegraphics<5>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16d}
      \end{column}
    \end{columns}
  \end{frame}

@@

  \section{Open issues}
  \begin{frame}{Open and unresolved issues}
  \setbeamercovered{transparent}
  \begin{enumerate}[<+(1) | invisible@-+>]
    \item Does $\Lambda_{div}$ unify the picture among the devices?
      \alert{No, does neutral pressure play a role?}
    \item If neutrals play a role is it at the divertor and/or at the midplane?
      \alert{Contradictory results if one include experiments with
        midplane puffing, JET and EIRENE simulations}
    \item Does L$_\parallel$ play any role? \alert{So far TCV suggests no
      dependence from f$_x$ scan but both TCV and MAST observe an
      I$_p$ dependence. MAST clearly state filament $\sigma_{\perp}$
      increases with L$_{\parallel}$ }
    \item Is cooling the divertor with fueling or seeding equivalent? \alert{Contradictory observations in AUG and JET}
    \item Do we observe same behavior in L and H-Mode? \alert{So
      far no as shown in H-Mode AUG. We need higher detachment
      condition and we need enough fueling}
      \item Is the density shoulder accompanied with an ion temperature shoulder?
%      \item What are the underlying mechanisms? 
    \end{enumerate}
  \end{frame}

@@

\begin{frame}{Focal points}
\begin{itemize}
	\item MST1 uniquely facilitates cross comparison between machines
	\item Several of the proposed experiments overlap and will be combined
	\item A cross machine experiment has the makings of settling open issues
	\item Therefore, we will allocate shots for cross-machine
          comparison but we must strive after similar machine configuration
	\item ....but also to other proposed experiments
	\item Cross machine L-mode experiments: 
	\begin{enumerate}
		\item Investigate the role of neutrals 
		\item $I_p$ and $q_{95}$ scans
	\end{enumerate}
	\item Cross machine H-mode experiments. 
\end{itemize}	
\end{frame}

\begin{frame}{L-mode experiments}
	\begin{itemize}
		\item The role of neutrals in the shoulder
                  formation is not understood
                  {\scriptsize(proponents: Carralero (\# 7), Militello
                    (\# 1, 2), Vianello (\# 13), Walkden (\# 9))}
		\begin{enumerate}
		\item We envisage to measure neutral gas profiles at the outboard midplane
                  using fast cameras with D$_{\alpha}$ filter, density
                  and temperature SOL profiles and appropriate code
                  (e.g. KN1D as in \parencite{Lipschultz:2005gg,Lipschultz:2016vk})
		\item Reciprocating probes and profiles available on
                  all machines
                \item Investigate role of fueling location (MAST)
		\end{enumerate}
		\item Disentangle the roles of $I_p$, $q_{95}$, and $L_{\|}$.
                  {\scriptsize(proponents: Carralero , Militello, Vianello, Tsui, Carralero (\# 7), Militello
                    (\# 1, 2), Vianello (\# 13), Walkden (\# 9), Tsui
                    (\# 15))}
		\begin{enumerate}
			\item Carry out parameter scans on all machines

@@

\begin{frame}{COMPASS collaboration}
\begin{itemize}
\item COMPASS is not an MST1 device, therefore no operational cost
  will be provided by MST1
\item However GA agrees that MST1 can provide part of the ppy for
  running experiments in line with MST1 topics
\item Proposal \# 14 is focused filaments dynamics and shoulder
  formation both in L and H-Mode
\item Experimental plan can be therefore agreed in order to perform
  similar experiments (e.g. I$_p$ and $q_{95}$ scan in L-Mode)
  ensuring proper comparison between devices
\item COMPASS suppose 3ppy for running and evaluation of experiment
  pertaining Topic-21, among Czeck and other european laboratories.
\end{itemize}
\end{frame}

\begin{frame}{There are urgent experiment which cannot be addressed}
\begin{itemize}
	\item Detailed ion temp measurements. Must eventually be
          completed in internal programs. Are you willing to share the
          information among Topic 21 even though done internally?
	\item Topology investigations? For example the comparison
          USN/LSN/DN is not included. Diagnostic capability of upper
          divertor is weaker than the lower one.
          Better to use reverse B$_t$  which is known
          to have influences in detachment as
          well \parencite{McLean}
	\item Strikepoint sweeping and cryo pump
	\item Parameters scans in H-mode
	\item X-point probes
\end{itemize}
\end{frame}

\begin{frame}{Modeling require}
  \begin{enumerate}
    \item Mean field modeling, in particular for neutral investigation.
      Available for AUG (EMC3-EIRENCE) and
      MAST-U (SOLPS). So far we lack investigation for these scenario
      targeting TCV. Manpower needed from edge and SOL modeling
    \item Turbulence code: we should need self-consistent interaction
      with neutrals. Possible tools \texttt{BOUT ++}, \texttt{HESEL},
      \texttt{GBS}, \texttt{Tokam-3X}.
      Some of them can be run by people who will, very
      likely, apply for this topic, some other not
  \end{enumerate}
\end{frame}

\begin{frame}{Future and Miscellaneous}
	\begin{itemize}
		\item Possibility to piggybag on
                  Quiescent/small ELM H-mode shots (Topic 5,6, 18)
		\item Shortly after this meeting we will organize
                  a follow-up meeting to reconcile ideas
                  from this discussion (date not fixed yet)
		\item All information will be gathered on the wiki
                \item Other channels for discussion/sharing. e.g. Slack
                  (mst1-topic21.slack.com), Github for code sharing, \ldots
		\item We aim to arrive to a proper shared database of
                  filaments between the machines, where properties
                  will be obtained using agreed approaches so that a
                  cross-machine comparison paper could at the end be
                  finalized at a certain point		
	\end{itemize}
\end{frame}

\begin{frame}{Discussion agenda}
	\begin{enumerate}
		\item Priorities. These are our personal ideas. Do we agree? 
		\item \underline{Common measurement techniques}. Filament
                  properties, $\Lambda_{div}$ estimate and location,
                  Neutrals, packing fraction, \ldots 
 	\end{enumerate}
\end{frame}
